\section{\module{sybasect} --- Interface to Sybase-CT library}

\declaremodule[sybasect]{extension}{sybasect}

\modulesynopsis{Interface to Sybase-CT library.}

This is not a complete reference to the Sybase CT library.  Sybase
produce excellent documentation which fully describes the use of the
CT library.

This section describes how to access the Sybase CT library while
using this wrapper module.

The \module{sybasect} extension contains the following:

\subsection{Types}

\begin{datadesc}{ContextType}
The type of \code{CS_CONTEXT} objects which wrap the Sybase
\code{CS_CONTEXT} structure pointer.
\end{datadesc}

\begin{datadesc}{ConnectionType}
The type of \code{CS_CONNECTION} objects which wrap the Sybase
\code{CS_CONNECTION} structure pointer.
\end{datadesc}

\begin{datadesc}{CommandType}
The type of \code{CS_COMMAND} objects which wrap the Sybase
\code{CS_COMMAND} structure pointer.
\end{datadesc}

\begin{datadesc}{BlkDescType}
The type of \code{CS_BLKDESC} objects which wrap the Sybase
\code{CS_BLKDESC} structure pointer.
\end{datadesc}

\begin{datadesc}{DataFmtType}
The type of \code{CS_DATAFMT} objects which wrap the Sybase
\code{CS_DATAFMT} structure.
\end{datadesc}

\begin{datadesc}{IODescType}
The type of \code{CS_IODESC} objects which wrap the Sybase
\code{CS_IODESC} structure.
\end{datadesc}

\begin{datadesc}{ClientMsgType}
The type of \code{CS_CLIENTMSG} objects which wrap the Sybase
\code{CS_CLIENTMSG} structure.
\end{datadesc}

\begin{datadesc}{ServerMsgType}
The type of \code{CS_SERVERMSG} objects which wrap the Sybase
\code{CS_SERVERMSG} structure.
\end{datadesc}

\begin{datadesc}{DataBufType}
The type of data buffers for sending and receiving data to and from
Sybase.  The type of object returned by \code{DataBuf('hello')}.
\end{datadesc}

\begin{datadesc}{NumericType}
The type used to store Sybase \code{CS_NUMERIC} and \code{CS_DECIMAL}
data values.
\end{datadesc}

\begin{datadesc}{DateTimeType}
The type used to store Sybase \code{CS_DATETIME} and
\code{CS_DATETIME4} data values.
\end{datadesc}

\begin{datadesc}{MoneyType}
The type used to store Sybase \code{CS_MONEY} and \code{CS_MONEY4}
data values.
\end{datadesc}

\subsection{Functions}

\begin{funcdesc}{set_global_ctx}{ctx}
The \module{sybasect} module uses a \code{CS_CONTEXT} structure
internally for conversions and calculations.  You must allocate a
context via \function{cs_ctx_alloc()} and establish the internal
context using this function.
\end{funcdesc}

\begin{funcdesc}{set_debug}{file}
Directs all debug text to the object passed in the \var{file}
argument.  The \var{file} argument must be either \code{None} or an
object which has \method{write(data)} and \method{flush()} methods.

The function returns the previous debug file.  The default file is
\code{None}.

Setting a debug file does not enable debug messages.  To produce debug
messages you need to set the \var{debug} member of a context,
connection, command, etc.
\end{funcdesc}

\begin{funcdesc}{cs_ctx_alloc}{\optional{version \code{= CS_VERSION_100}}}
Calls the Sybase-CT \function{cs_ctx_alloc()} function:

\begin{verbatim}
result = cs_ctx_alloc(version, &ctx);
\end{verbatim}

Returns a tuple containing the Sybase result code and a new instance
of the \class{CS_CONTEXT} class constructed from the \var{ctx} value
returned by \function{cs_ctx_alloc()}.  \code{None} is returned as the
\class{CS_CONTEXT} object if the result code is not \code{CS_SUCCEED}.
\end{funcdesc}

\begin{funcdesc}{cs_ctx_global}{\optional{version \code{= CS_VERSION_100}}}
Calls the Sybase-CT \function{cs_ctx_global()} function:

\begin{verbatim}
result = cs_ctx_global(version, &ctx);
\end{verbatim}

Returns a tuple containing the Sybase result code and a new instance
of the \class{CS_CONTEXT} class constructed from the \var{ctx} value
returned by \function{cs_ctx_global()}.  \code{None} is returned as the
\class{CS_CONTEXT} object if the result code is not \code{CS_SUCCEED}.
\end{funcdesc}

\begin{funcdesc}{DataBuf}{obj}
Return a new instance of the \class{DataBuf} class.  The \var{obj}
argument is used to initialise the \class{DataBuf} object.

For all types of \var{obj} other than \class{CS_DATAFMT} a buffer will
be initialised which contains a single value.

When \var{obj} is a \class{CS_DATAFMT} object an empty buffer will be
created according to the attributes of the \class{CS_DATAFMT} object.
It is most common to create and bind a buffer in a single operation
via the \method{ct_bind()} method of the \class{CS_COMMAND} class.

For example, the following code creates a set of buffers for
retrieving 16 rows at a time.  Note that it is your responsibility to
ensure that the buffers are not released until they are no longer
required.

\begin{verbatim}
status, num_cols = cmd.ct_res_info(CS_NUMDATA)
if status != CS_SUCCEED:
    raise 'ct_res_info'
bufs = []
for i in range(num_cols):
    status, fmt = cmd.ct_describe(i + 1)
    if status != CS_SUCCEED:
        raise 'ct_describe'
    fmt.count = 16
    status, buf = cmd.ct_bind(i + 1, fmt)
    if status != CS_SUCCEED:
        raise 'ct_bind'
    bufs.append(buf)
\end{verbatim}
\end{funcdesc}

\begin{funcdesc}{numeric}{obj \optional{, precision \code{= -1}} \optional{, scale \code{= -1}}}
Return a new instance of the \class{Numeric} class.

The \var{obj} argument can accept any of the following types;
\code{IntType}, \code{LongType}, \code{FloatType}, \code{StringType},
or \code{NumericType}.

If greater than or equal to zero the \var{precision} and \var{scale}
arguments are used as the \member{precision} and \member{scale}
attributes of the \code{CS_DATAFMT} which is used in the Sybase
\function{cs_convert()} function to create the new \class{NumericType}
object.  The default values for these arguments depends upon the type
being converted to \class{NumericType}.

\begin{longtable}{l|l|l}
Type & \var{precision} & \var{scale} \\
\hline
\code{IntType}     & \code{16} & \code{0} \\
\code{LongType}    & \# of digits in \code{str()} conversion & \code{0} \\
\code{FloatType}   & \code{CS_MAX_PREC} & \code{12} \\
\code{StringType}  & \# digits before decimal point & \# digits after decimal point \\
\code{NumericType} & input precision & input scale \\
\end{longtable}
\end{funcdesc}

\begin{funcdesc}{money}{obj \optional{, type \code{= CS_MONEY_TYPE}}}
Return a new instance of the \class{Money} class.

The \var{obj} argument can accept any of the following types;
\code{IntType}, \code{LongType}, \code{FloatType}, \code{StringType},
or \code{MoneyType}.

Passing \code{CS_MONEY4_TYPE} in the \var{type} argument will create a
\code{smallmoney} value instead of the default \code{money}.
\end{funcdesc}

\begin{funcdesc}{datetime}{str \optional{, type \code{= CS_DATETIME_TYPE}}}
Return a new instance of the \class{DateTime} class.

The \var{str} argument must be a string.

Passing \code{CS_DATETIME4_TYPE} in the \var{type} argument will
create a \code{smalldatetime} value instead of the default
\code{datetime}.
\end{funcdesc}

\begin{funcdesc}{sizeof_type}{type_code}
Returns the size of the type identified by the Sybase type code
specified in the \var{type_code} argument.

The values expected for the \var{type_code} argument things like;
\code{CS_CHAR_TYPE}, \code{CS_TINYINT_TYPE}, etc.
\end{funcdesc}

\begin{funcdesc}{CS_DATAFMT}{}
Return a new instance of the \class{CS_DATAFMT} class.  This is used
to wrap the Sybase \code{CS_DATAFMT} structure.
\end{funcdesc}

\begin{funcdesc}{CS_IODESC}{}
Return a new instance of the \class{CS_IODESC} class.  This is used
to wrap the Sybase \code{CS_IODESC} structure.
\end{funcdesc}

\begin{funcdesc}{CS_LAYER}{msgnumber}
Return the result of applying the Sybase \code{CS_LAYER} macro to the
\var{msgnumber} argument.
\end{funcdesc}

\begin{funcdesc}{CS_ORIGIN}{msgnumber}
Return the result of applying the Sybase \code{CS_ORIGIN} macro to the
\var{msgnumber} argument.
\end{funcdesc}

\begin{funcdesc}{CS_SEVERITY}{msgnumber}
Return the result of applying the Sybase \code{CS_SEVERITY} macro to the
\var{msgnumber} argument.
\end{funcdesc}

\begin{funcdesc}{CS_NUMBER}{msgnumber}
Return the result of applying the Sybase \code{CS_NUMBER} macro to the
\var{msgnumber} argument.
\end{funcdesc}

\input{ctxobj}
\input{localeobj}
\input{connobj}
\input{cmdobj}
\input{msgsobj}
\input{datafmtobj}
\input{databufobj}
\input{iodescobj}
\input{blkdescobj}
